---
title: "Renv"
author: "Rockefeller University, Bioinformatics Resource Center"
date: "https://rockefelleruniversity.github.io/RU_reproducibleR/"
output: 
  html_document:
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
params:
  toMessage: false    
---


```{r, setup,include=FALSE}
toMessage <- ifelse(!is.null(params$toMessage),params$toMessage,FALSE)
message(toMessage)
```

<p>&nbsp;</p>
These exercises follow the material in [Reproducible R](https://rockefelleruniversity.github.io/RU_reproducibleR//). 

**Exercise 1** - **Packages and R**

- What version of R are you using?
```{r, eval=FALSE,echo=toMessage}
sessionInfo()
```

- What version of ggplot2 do you have?
```{r, eval=FALSE,echo=toMessage}
sessionInfo("ggplot2")
```

- What version of Bioconductor are you using?

```{r, eval=FALSE,echo=toMessage}
library(BiocManager)
version()
```

- Load in Rmarkdown, ggplot2 and plotly. How many packages do  I have loaded in now?
```{r, eval=FALSE,echo=toMessage}
library(rmarkdown)
library(ggplot2)
library(plotly)
sessionInfo()
```

**Exercise 2** - **Renv**

- Create a new project
- Initialize a Renv for this project
- *scriptToConvertToRMarkdown.r* into this project
- Update the status and create a snapshot
- Create another new project and add all the essential files to this new project. Use Renv to restore the environment in the new project. 



