---
title: "Reproducible R"
author: "Rockefeller University, Bioinformatics Resource Center"
date: "https://rockefelleruniversity.github.io/RU_reproducibleR/"
output: 
  html_document:
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
params:
  toMessage: false    
---

```{r, setup,include=FALSE}
toMessage <- ifelse(!is.null(params$toMessage),params$toMessage,FALSE)
message(toMessage)
```

<p>&nbsp;</p>
These exercises follow the material in [Reproducible R](https://rockefelleruniversity.github.io/RU_reproducibleR//). 

**Exercise 1** - **Rmarkdown basics**

- Open up markdownExampleDefaultStyles.Rmd and compare each section to the rendered output: [Example HTML Default style.](../../scripts/markdownExampleDefaultStyles.html). Review how each piece of markdown controls the final rendered document. There are several examples we have not covered in the slides. 

- Open up markdownExample.Rmd in the scripts directory and compare to the rendered output: [Example HTML with extra style.](../../scripts/markdownExample.html). What has changed?
```{r, eval=FALSE,echo=toMessage}
# Different Styles in YAML
# Updated table of contents in YAML
# Set Up chunk - we can use this to control defaults for code chunk arguments. In this case we have turned echo off for every chunk. 
```

**Exercise 2** - **Convert a Rmarkdown**

- Open scriptToConvertToRMarkdown.r in scripts directory and save as new name.
- Convert this script to an Rmarkdown document inside RStudio.
- Update YAML to reflect author, date, title etc.
- Break the code down into chunks. Add some code chunk arguments to control figure sizes and/or echo of code.
- Add appropriate annotation and session information.
- Try making both interactive and non-interactive versions of the tables and plots.
- Render the document.

**Exercise 3** - **Make a Rmarkdown**
- Make your own Rmarkdown document from scratch
- Include YAML parameters like date
- Read in file "readThisTable.csv"
- Run some summary stats and plots
- Use chunk arguments: fig.height/fig.width/fig.path/dev to save plots in pdf and png format

**Exercise 4** - **Quarto**
- Convert your Rmd from exercise 2 into a Quarto document
- Keep the parameters the same. Render and review the result. Compare it to the Rmd output. 
- Try and update all code arguments

**Additional Exercise**
- Take some of your own scripts. They can either be python or R. Try and convert them into an Rmarkdown/Quarto format so you can create a nice report summary of that piece of work.
