---
title: "Conda and Herper<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>"
author: "Rockefeller University, Bioinformatics Resource Centre"
date: "https://rockefelleruniversity.github.io/RU_reproducibleR/"
output: 
  xaringan::moon_reader:
    css: ["default", "metropolisCustom.css", "metropolis-fontsCustom.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
  html_document:
    toc: true # table of content true
    toc_float: yes
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
params:
  isSlides: "no"
---
```{r setup, include=FALSE}
suppressPackageStartupMessages(require(knitr))
knitr::opts_chunk$set(echo = TRUE, tidy = T, eval=F)
```

```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides != "yes"){
  cat("# Conda

---
"    
  )
  
}

```


## What is Conda?

(Conda)[https://conda.org/] is an open-source and cross-platform package and environment management system. 

It was originally developed to help manage packages for Python, but that has expanded beyond python and is very popular with R users as well. 

---
## How does Conda work?

---
## Why Conda?

* Simplify Package Management
* Cross-Platform Compatibility
* Extensive Package Ecosystem
* Environment Management

---
## Conda vs miniconda vs ...

There are many different repository tools in the conda ecosystem. They all have key differences. 

Conda - The core package manager
Miniconda - A minimal install of Conda
Anaconda - A maximal install of Conda
Mamba - A reimplementation of Conda built in C++ [FASTER]
Minimamba - A minimal install of Mamba

---
## Working with Conda

* Work with miniconda (or Herper)
* Always install software into new environments (never the base)



---
```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides == "yes"){
  cat("class: inverse, center, middle

# Herper

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---
"    
  )
}else{
  cat("# Herper

---
"    
  )
  
}

```

## What is Herper

Herper is an R package that provides a simple toolset to install and manage Conda packages and environments from within the R console.

It is built up from the package *reticulate* which is used to run python from within R.

---
## Installing Tools

The **install_CondaTools()** function allows the user to specify required Conda software and the desired environment to install into.

Miniconda is installed as part of the process (by default into the r-reticulate's default Conda location - `r reticulate::miniconda_path()`) and the user's requested conda environment built within the same directory (by default `r file.path(reticulate::miniconda_path(),"envs","USERS_ENVIRONMENT_HERE")`). 

```{r}
library(Herper)
install_CondaTools(tools = "star", env = "rnaseq")

```

---
## Installing Tools with speicific conda

If you already have Miniconda installed or you would like to install to a custom location, you can specify the path with the *pathToMiniConda* parameter.

```{r}
my_miniconda <- "~/Desktop/My_Conda"

install_CondaTools(tools = "star", env = "rnaseq", pathToMiniConda = my_miniconda)

```

---
## Adding more tools

It's easy to add more tools. You just use *install_CondaTools()* again, but an extra argument also needs to be added: *updateEnv = TRUE*.

By default when we run this command, the path to Conda and the environment is returned. We can save this as a variable for later. 

```{r}
conda_paths <- install_CondaTools(tools = c("salmon","kallisto"), env = "rnaseq", updateEnv = TRUE , pathToMiniConda = my_miniconda)

conda_paths

```


---
## Checking for packages in an environment with **list_CondaPkgs**

The **list_CondaPkgs** function allows users to check what packages are installed in a given environment.

```{r list_CondaPkgs}
list_CondaPkgs("rnaseq", pathToMiniConda = my_miniconda)
```


---
## Finding Conda packages with **conda_search**

If the user is unsure of the exact name, or version of a tool available on conda, they can use the **conda_search** function. Searches will find close matches for incorrect queries.

```{r conda_search}
conda_search("salmo", pathToMiniConda = my_miniconda)
```

---
## Finding Conda package versions with **conda_search**

If you have the exact name you can search for what versions are available on Conda. 

```{r conda_search2}
conda_search("salmon", pathToMiniConda = my_miniconda)
```

---
## Finding Conda package versions with **conda_search**

Specific package versions can be searched for using the Conda format i.e. "salmon==1.3", "salmon>=1.3" or "salmon<=1.3".

```{r conda_search_nuance}
conda_search("salmon<=1.0", pathToMiniConda = my_miniconda)

```

---
## Installing specific versions

We can use the same version nomenclature to also install these tools. Here we will downgrade the version of salmon we have installed. 

```{r}
conda_paths <- install_CondaTools(tools = "salmon<=1.0", env = "rnaseq", updateEnv = T, pathToMiniConda = my_miniconda)

```

---
## Installing specific versions

We can now see that we have downgraded 
```{r}
library(magrittr)
library(dplyr)
list_CondaPkgs("rnaseq", pathToMiniConda = my_miniconda) %>% dplyr::filter(name=="salmon")
```

---
## Running software

Once installed within a conda environment, many external software can be executed directly from the conda environment's bin directory without having to perform any additional actions. 

The Herper package uses the **[withr](https://withr.r-lib.org)** family of functions (**with_CondaEnv()** and **local_CondaEnv()**) to provide methods to **temporarily** alter the system PATH and to add or update any required environmental variables. This is done without formally activating your environment or initializing your conda.

---
## **with_CondaEnv**

The **with_CondaEnv** allows users to run R code with the required PATH and environmental variables automatically set. The **with_CondaEnv** function simply requires the name of conda environment and the code to be executed within this environment. Additionally we can also the **pathToMiniconda** argument to specify any custom miniconda install location. 

Here we use the **with_CondaEnv** to allow us to *see* the tools installed in out our "rnaseq" environment. We can then use the R function *system2()* to run some terminal/command line code from within R. In this case we want to run the salmon help. 

```{r with_condaenv_SalmonWithCondaEnvEval,echo=TRUE,eval=FALSE,tidy=FALSE}
res <- with_CondaEnv("rnaseq",
                      system2(command="salmon",args = "help",stdout = TRUE),
                      pathToMiniConda=my_miniconda)
res
```

---
## **local_CondaEnv**

The **local_CondaEnv** function acts in a similar fashion to the **with_CondaEnv** function: it will update the required PATH and environmental variable so you can access the tools you need. The PATH and environmental variables will persist though until the current function ends.

**local_CondaEnv** is best used within a user-created function, allowing access to the Conda environment's PATH and variables from within the the function itself but resetting all environmental variables once complete.

```{r with_condaenv_SalmonLocalCondaEnv,echo=TRUE,eval=FALSE}
salmonHelp<- function(){
  local_CondaEnv("rnaseq", pathToMiniConda=my_miniconda)
  helpMessage <- system2(command="salmon",args = "help",stdout = TRUE)
  helpMessage
}
salmonHelp()
```


---
## A practical example

To further demonstrate this we will use the first command from the [seqCNA](https://www.bioconductor.org/packages/release/bioc/html/seqCNA.html) vignette. This step requires samtools. If this is not installed and available there is an error. 
```{r with_condaenv_R}
library(seqCNA)
data(seqsumm_HCC1143)

try(rco <- readSeqsumm(tumour.data = seqsumm_HCC1143), silent = FALSE)
```

Samtools is listed as a System Requirement for seqCNA, so we can first use **install_CondaSysReqs()** to install samtools. In this case we are installing samtools in the environment: seqCNA_env. We can then run the seqCNA command using **with_CondaEnv** specifying that we want to use our environment containing samtools. seqCNA can then find samtools and execute successfully. 

```{r, echo=T, eval=F}
install_CondaSysReqs(pkg="seqCNA",env="seqCNA_env",pathToMiniConda=my_miniconda)
rco <- with_CondaEnv(new="seqCNA_env",readSeqsumm(tumour.data=seqsumm_HCC1143)
 ,pathToMiniConda = myMiniconda)
summary(rco)
``` 

---
## Reproducible environments

Once you are done you have an environment that is functioning well you will want to save it. One way to back it up is to export a snapshot. For Conda this snapshot is a *.yml* file. These files contain all information about the environment you would need in order to rebuild or share it for collaboration.


```{r export}
yml_name <- paste0("rnaseq_", format(Sys.Date(), "%Y%m%d"), ".yml")
export_CondaEnv("rnaseq", yml_name, pathToMiniConda = my_miniconda)
```

---
## Importing from *.yml*

The **import_CondaEnv** function allows the user to create a new conda environment from a *.yml* file. These can be previously exported from **export_CondaEnv**, conda, renv or manually created. 

Users can simply provide a path to the YAML file for import. They can also specify the environment name, but by default the name will be taken from the YAML. 

```{r import, eval=F}
testYML <- system.file("extdata/test.yml", package="Herper")
import_CondaEnv(yml_import=testYML, pathToMiniConda = my_miniconda)
```

---
## Other tricks
- Mamba - Mamba 
- Verbosity
- list conda env

---
```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides == "yes"){
  cat("class: inverse, center, middle

# Pip

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---
"    
  )
}else{
  cat("# Pip

---
"    
  )
  
}

```


## What about Pip?

Pip is an alternative system to install python tools. But it typically only works with python tools.

Pip doesn't handle conflicting dependencies, and will upgrade/downgrade software without checking if something else depends on it. 

Even if this doesn't break your tools, it may give different results and hinder reproducibility as unsupported versions of tools could be used.

Conda is much more considered, and checks installs to make sure that all your dependencies do not conflict. 

---
## Why use Pip?

* Not very Conda build is perfect
* Not every tool on conda

Sometimes we are forced to use pip. But that does not mean we have to leave our environment behind. 

---
## Export env

```{r}

list_CondaPkgs("rnaseq", pathToMiniConda = my_miniconda)
```





---
## Pip install

```{r}
install_CondaTools("pip", "rnaseq", pathToMiniConda = my_miniconda, updateEnv = TRUE)

with_CondaEnv("rnaseq",
                      system2(command="pip",args = c("install", "scanpy"),stdout = TRUE),
                      pathToMiniConda=my_miniconda)


with_CondaEnv("rnaseq",
                      system2(command=paste0(conda_paths$pathToEnvBin,"/pip"),args = c("install", "scanpy"),stdout = TRUE),
                      pathToMiniConda=my_miniconda)

```

---
## Check envs

```{R}

list_CondaPkgs("rnaseq", pathToMiniConda = my_miniconda)
```


---
```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides == "yes"){
  cat("class: inverse, center, middle

# Renv and Conda

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---
"    
  )
}else{
  cat("# Renv and Conda

---
"    
  )
  
}

```




```{r}
renv::use_python(python = "/path/to/python")

```




```{r}

renv::snapshot() calls conda env export > environment.yml to save the set of installed Python packages;

renv::restore() calls conda env [create/update] --file environment.yml to install the previously-recorded set of Python packages

```





---
## Further Resources


 

---
## Exercises

Exercise on Reproducibility in R can be found [here](../../exercises/exercises/ReproducibleR_exercise.html)


---
## Contact

Any suggestions, comments, edits or questions (about content or the slides themselves) please reach out to our [GitHub](https://github.com/RockefellerUniversity/Reproducible_R/issues) and raise an issue.


