---
title: "Environments and Renv<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>"
author: "Rockefeller University, Bioinformatics Resource Centre"
date: "https://rockefelleruniversity.github.io/RU_reproducibleR/"
output: 
  xaringan::moon_reader:
    css: ["default", "metropolisCustom.css", "metropolis-fontsCustom.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
  html_document:
    toc: true # table of content true
    toc_float: yes
    depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
params:
  isSlides: "no"
---
```{r setup, include=FALSE}
suppressPackageStartupMessages(require(knitr))
knitr::opts_chunk$set(echo = TRUE, tidy = T)
```

```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides != "yes"){
  cat("# Environments

---
"    
  )
  
}

```

## Set Up

All prerequisites, links to material and slides for this course can be found on github.

* [Reproducible_R](https://rockefelleruniversity.github.io/Reproducible_R/)

Or can be downloaded as a zip archive from here.

* [Download zip](https://github.com/rockefelleruniversity/Reproducible_R/zipball/master)

---
## Course materials

Once the zip file in unarchived. All presentations as HTML slides and pages, their R code and HTML practical sheets will be available in the directories underneath.

* **presentations/slides/**
Presentations as an HTML slide show.
* **presentations/singlepage/** 
Presentations as an HTML single page.
* **presentations/r_code/**
R code in presentations.
* **exercises/**
Practicals as HTML pages. 
* **answers/**
Practicals with answers as HTML pages and R code solutions. 


---
## Set the Working directory

Before running any of the code in the practicals or slides we need to set the working directory to the folder we unarchived. 

You may navigate to the unarchived Reproducible_R folder in the Rstudio menu.

**Session -> Set Working Directory -> Choose Directory**

or in the console.

```{r setwd_introtoR,eval=F} 
setwd("/PathToMyDownload/Reproducible_R-master/r_course")
# e.g. setwd("~/Downloads/Reproducible_R-master/r_course")
```

---
```{r, results='asis',include=TRUE,echo=FALSE}
if(params$isSlides == "yes"){
  cat("class: inverse, center, middle

# What are containers?

<html><div style='float:left'></div><hr color='#EB811B' size=1px width=720px></html> 

---
"    
  )
}else{
  cat("# What are containers?

---
"    
  )
  
}

```













---
## Further Resources


 

---
## Exercises

Exercise on Reproducibility in R can be found [here](../../exercises/exercises/ReproducibleR_exercise.html)


---
## Contact

Any suggestions, comments, edits or questions (about content or the slides themselves) please reach out to our [GitHub](https://github.com/RockefellerUniversity/Reproducible_R/issues) and raise an issue.


